---
title: "Shiny"
output: ioslides_presentation
runtime: shiny
---

<!-- 
author: "Shiwei Lan"
title: "Shiny"
date:  "`r format(Sys.time(), '%B %d, %Y')`  \n"
-->

<style>
.gdbar img {
  width: 150px !important;
  height: 80px !important;
  margin: 8px 8px;
}

.gdbar {
  width: 200px !important;
  height: 100px !important;
}

</style>

<style type="text/css">
body, p, div, h1, h2, h3, h4, h5, ul {
  color: black;
  font-family: Modern Computer Roman;
}
h2 {
color: #8C1D40;
}

</style>


<style type="text/css">
slides > slide:not(.nobackground):after {
  content: '';
}
</style>

<style type="text/css">
slides > slide {
  background: linear-gradient(#ffffff, #ffffff 85%, #ffffff);
  background-color: white;
  }
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Shiny Presentation {.smaller}

This presentation illustrates  interactive slides with two embedded Shiny apps. First app is included in the template when you create a new RMarkdown file by clicking on 

`New File -> R Markdown... -> Shiny -> Shiny Presentation`. 

The other one is `t_vs_Z` app. 

For more info on shiny apps, go to <http://rmarkdown.rstudio.com/authoring_shiny.html>


## `faithful` dataset {.smaller}

The data frame `faithful` from the `datasets` built-in package contains the data of waiting times between eruptions and the duration of the eruptions for the Old Faithful geyser in Yellowstone National Park, Wyoming.

```{r}
head(faithful)
```



## Interactive Plot; Old Faithful Eruptions {.smaller}

```{r eruptions, out.width="50%"}
inputPanel(
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  
  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Old Faithful Geyser eruption duration")
  
  dens <- density(faithful$eruptions, adjust = input$bw_adjust, kernel = "gaussian")
  lines(dens, col = "blue")
})
```


## Code for Old Faithful Geyser Eruptions{.smaller}

```{r, echo=T, eval=F}
inputPanel(
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  
  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser eruption duration")
  
  dens <- density(faithful$eruptions, adjust = input$bw_adjust, kernel="gaussian")
  lines(dens, col = "blue")
})
```

The actual bandwidth used is in fact `adjust * bw`. This makes it easy to specify values like ‘half the default’ bandwidth. The default bandwidth is 1 (with the default `bw="nrd0"` and `adjust=1`)


## {}

```{r, echo=F}

library(latex2exp)

shinyApp(
  
  ui = fluidPage(

    # Application title
    titlePanel(h5("Densities of Standard Normal and Student t-distribution")),

    # Sidebar with a slider input for number of bins
    sidebarLayout(
        sidebarPanel(
            sliderInput("d",
                        "Degree of freedom: d",
                        min = 1,
                        max = 50,
                        step = 1,
                        value = 1),
            sliderInput("a",
                        "tail @: a",
                        min = 0,
                        max = 5,
                        step = 0.01,
                        value = 3),
            checkboxInput(inputId="show_t_area", 
                          label="show P(t_d > a) area", 
                          value = FALSE),
            checkboxInput(inputId="show_z_area", 
                          label="show P(Z > a) area", 
                          value = FALSE)
        ),

        # Show a plot of the generated distribution
        mainPanel(
            plotOutput("distPlot", height = "500px")
        )
    )
),  ## and of ui and its fluidPage

server = function(input, output) {

    output$distPlot <- renderPlot({

        # generate bins based on input$bins from ui.R
        xmax = max(4,input$a)
        xvals = seq(-xmax, xmax,by=0.01)
       
        plot(xvals, dnorm(xvals, mean=0, sd=1),
             xlab = "x", ylab = "y",
             xlim = c(-xmax, xmax),
             type="l", col="red", 
             main=TeX("Densities of $Z$ and $t_d$"))
        lines(xvals, dt(xvals,df=input$d), col="blue")
        abline(v=input$a,lty="dashed")
        abline(h=0, lty="solid")
        xtick = c(seq(-4,4,by=2),input$a)
        axis(side=1, at=xtick, labels = c(seq(-4,4,by=2),"a"))
        legend("topleft", 
               legend=c("Z",TeX(paste("$t_d$ ;   d = ",as.character(input$d)))),
               col=c("red","blue"), lty="solid",bty="n",
               text.col = c("red","blue"))
        text(input$a-0.2, 0.3, 
             labels = TeX(paste("a = ", as.character(input$a), sep="")),
             adj=0)  ## adj=0 means text is left-aligned
        
        text(-4.2,0.33, 
             labels=TeX(paste("P(t_d > a) = ",
                              as.character(
                                  sprintf(pt(q=input$a,df=input$d,lower.tail=FALSE),fmt = '%#.7f')))),
             adj=0, col="blue")  ## adj=0 means text is left-aligned
        text(-4.2,0.28, 
             labels=TeX(paste("P( Z > a) = ",
                              as.character(
                                  sprintf(pnorm(q=input$a,lower.tail=FALSE),fmt = '%#.7f')))),
             adj=0, col="red") ## adj=0 means text is left-aligned
        
            
            xnew = seq(input$a,xmax,by=0.01)

            if (input$show_t_area == TRUE){            
            ##plotting area P(t_n>a)
            polygon(x=c(input$a,xnew,xnew[length(xnew)]), 
                    y=c(0,dt(xnew,df=input$d),0), 
                    density=15, col="blue")
            }
            if (input$show_z_area == TRUE){
            ##plotting area P(Z>a)
            polygon(x=c(input$a,xnew,xnew[length(xnew)]), 
                    y=c(0,dnorm(xnew,mean=0,sd=1),0), 
                    angle=135, density=40, col="red")
        }
        
    })

}, ## end of the server function

  options = list(height = "100%")

)  ##end of shinyApp() function
```


