---
title:  Lab 2 SOLUTION 
author: "Grader"
date: "3/25/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "", out.width="0.5 \\textwidth", fig.align = "center")
```

``` {r, eval = TRUE, echo = FALSE, message=FALSE, warning=FALSE}
library(ggplot2movies)
library(dplyr)
library(ggplot2)
```


## Question 1

**What is the range of years of production of the movies of this data set?**

```{r}
data(movies)
range(movies$year)
```

The production years of the oldest and most recent movie are 1893 and 2005 respectively.


## Question 2

**What proportion of movies have their budget included in this data base, and what proportion doesn’t?
What are top 5 most expensive movies in this data set?**

```{r}
nobug = sum(is.na(movies$budget))
N = dim(movies)[1]
prop = (N-nobug)/N
print(prop)
```

The proportion of movies have their budget included in this data is $8.87\%$, and the oppersite proportion is $91.13\%$.

The top 5 most expensive movies are

```{r}
movies %>% top_n(5,budget) %>% select(title)
```

## Question 3

**What are top 5 longest movies?**

```{r}
movies %>% top_n(5,length) %>% select(title,length)
```

## Question 4

** Of all short movies, which one is the shortest? Which one is the longest? How long are the
shortest and the longest short movies? **

The longest shore movie is
```{r}
movies %>% filter(Short==1) %>% top_n(1,length) %>% select(title,length)
```

The shortest movies are
```{r}
movies %>% filter(Short==1) %>% top_n(-1,length) %>% select(title,length)
```

## Question 5

**How many movies of each genre (action, animation, comedy, drama, documentary, romance, short) are there in this data base?**

```{r}
tep=movies %>% select(Action:Short) %>% colSums 
data.frame(genre = names(tep),count = tep) %>% ggplot(aes(x=genre,y=count,fill=genre)) +
  geom_bar(stat="identity") + theme(legend.position = "none") +
  geom_text(aes(label=count),position=position_dodge(width=0.9), vjust=-0.25)
```

## Question 6

**What is the average rating of all movies within each genre? **

```{r}
movies %>% tidyr::gather(key="genre",value="identity","Action","Animation","Comedy","Drama",
                  "Documentary", "Romance", "Short") %>%
  filter(identity == 1) %>%
  group_by(genre)  %>%
  summarise(average = mean(rating)) %>% ggplot(aes(x=genre,y=average,fill=genre)) +
  geom_bar(stat="identity") + theme(legend.position = "none") +
  geom_text(aes(label=round(average,2)),position=position_dodge(width=0.9), vjust=-0.25)
```

## Question 7

**What is the average rating of all movies within each genre that were produced in the years 2000-2005? **

```{r}
movies %>% tidyr::gather(key="genre",value="identity","Action","Animation","Comedy","Drama",
                  "Documentary", "Romance", "Short") %>%
  filter(identity == 1) %>%
  filter(year >= 2000 & year < 2005) %>% 
  group_by(genre)  %>%
  summarise(average = mean(rating)) %>% ggplot(aes(x=genre,y=average,fill=genre)) +
  geom_bar(stat="identity") + theme(legend.position = "none") +
  geom_text(aes(label=round(average,2)),position=position_dodge(width=0.9), vjust=-0.25)
```

## Question 8

**For each of the first 6 genres (not including short movies) consider only movies from 1990 until the last year recorded and plot a function of the number of movies in this data base of corresponding genre produced by year, for years from 1990 until the last year recorded.**

```{r}
movies %>% 
  tidyr::gather(key = "genre", value =  "identity", "Action", "Animation", "Comedy", 
         "Drama", "Documentary", "Romance") %>% 
  filter(identity == 1) %>%
  filter (year >= 1990) %>%
  group_by(genre, year) %>% 
  add_tally(name = "Movies_per_Year") %>%
  group_by(year, genre, Movies_per_Year) %>% 
  summarize() %>%
  ggplot(aes(x = year, y = Movies_per_Year, colour = genre)) + 
  geom_line(size=1.05) + geom_point() + ggtitle('Total Movies Produced by Genre from 1990-2005')
```

